---
title: "Neural Network Analysis"
author: "Ahsan Ahmad, Varun Selvam, Nikita Muddapati"
date: "2024-11-22"
output: 
  html_document:
    theme: flatly
    df_print: paged
    toc: true
    toc_float:
      smooth_scroll: true
execute:
  include: true
  eval: true    
  warning: false
  message: false
---

# Q1: From what perspective are you conducting the analysis? (Who are you? / Who are you working for?)

This analysis is conducted from the perspective of a real-estate analytics team at Zillow. Our objective is to create a robust neural network model to predict house prices in the suburbs of Ames, Iowa, based on historical housing data.

# Q2: What is your question?

We aim to build a neural network model utilizing the avNNET and RWeka R packages to predict house prices in Ames, Iowa. The goal is to identify key predictors of house prices and achieve high accuracy and generalizability across training and test datasets.

# Q3: Describe your dataset(s) including URL (if available).

The data-set that we are using is from the Houses Prices- Advanced Regression Techniques. This data-set has 1460 observations and 80 predictors that describe various aspects of the houses in Ames, Iowa. However after cleaning, the dataset has 1383 observtions and 78 predictors. Please see the Cleaning Data Section for more details about how the data cleaning was conducted.

URL: https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data

# Q4: What are your independent and dependent variables? Include variable type.

*Numeric Variables:*

MSSubClass: The building class
LotFrontage: Linear feet of street connected to property
LotArea: Lot size in square feet
YearBuilt: Original construction date
YearRemodAdd: Remodel date
MasVnrArea: Masonry veneer area in square feet
BsmtFinSF1: Type 1 finished square feet
BsmtFinSF2: Type 2 finished square feet
BsmtUnfSF: Unfinished square feet of basement area
TotalBsmtSF: Total square feet of basement area
1stFlrSF: First Floor square feet
2ndFlrSF: Second floor square feet
LowQualFinSF: Low quality finished square feet (all floors)
GrLivArea: Above grade (ground) living area square feet
BsmtFullBath: Basement full bathrooms
BsmtHalfBath: Basement half bathrooms
FullBath: Full bathrooms above grade
HalfBath: Half baths above grade
Bedroom: Number of bedrooms above basement level
Kitchen: Number of kitchens
TotRmsAbvGrd: Total rooms above grade (does not include bathrooms)
Fireplaces: Number of fireplaces
GarageYrBlt: Year garage was built
GarageCars: Size of garage in car capacity
GarageArea: Size of garage in square feet
WoodDeckSF: Wood deck area in square feet
OpenPorchSF: Open porch area in square feet
EnclosedPorch: Enclosed porch area in square feet
3SsnPorch: Three season porch area in square feet
ScreenPorch: Screen porch area in square feet
PoolArea: Pool area in square feet
MiscVal: Value of miscellaneous feature
MoSold: Month Sold
YrSold: Year Sold
*Categorical Variables:*

MSZoning: The general zoning classification
Alley: Type of alley access
LotShape: General shape of property
LandContour: Flatness of the property
LotConfig: Lot configuration
LandSlope: Slope of property
Neighborhood: Physical locations within Ames city limits
Condition1: Proximity to main road or railroad
Condition2: Proximity to main road or railroad (if a second is present)
BldgType: Type of dwelling
HouseStyle: Style of dwelling
OverallQual: Overall material and finish quality
OverallCond: Overall condition rating
RoofStyle: Type of roof
RoofMatl: Roof material
Exterior1st: Exterior covering on house
Exterior2nd: Exterior covering on house (if more than one material)
MasVnrType: Masonry veneer type
ExterQual: Exterior material quality
ExterCond: Present condition of the material on the exterior
Foundation: Type of foundation
BsmtQual: Height of the basement
BsmtCond: General condition of the basement
BsmtExposure: Walkout or garden level basement walls
BsmtFinType1: Quality of basement finished area
BsmtFinType2: Quality of second finished area (if present)
Heating: Type of heating
HeatingQC: Heating quality and condition
Electrical: Electrical system
KitchenQual: Kitchen quality
Functional: Home functionality rating
FireplaceQu: Fireplace quality
GarageType: Garage location
GarageFinish: Interior finish of the garage
GarageQual: Garage quality
GarageCond: Garage condition
PavedDrive: Paved driveway
PoolQC: Pool quality
Fence: Fence quality
MiscFeature: Miscellaneous feature not covered in other categories
SaleType: Type of sale
SaleCondition: Condition of sale
*Binary Variables:*

Street: Type of road access
Utilities: Type of utilities available
CentralAir: Central air conditioning

# Q5: How are your variables suitable for your analysis method?

Neural networks can model complex, non-linear relationships between predictors and the target variable. The dataset's combination of numeric and categorical variables makes it well-suited for neural networks, as these models can effectively learn patterns and interactions across multiple data types.

The dataset was preprocessed as follows to ensure compatibility with the avNNET and RWeka packages:

Numeric Variables: Scaled to a range of [0, 1] for improved convergence.
Categorical Variables: One-hot encoded.
Missing Data: Imputed using median/mode imputation.

# Q6: What are your conclusions?
The final neural network model was trained using grid search (RWeka Package) and evaluated on a held-out test set. Key findings include:

*Model Performance:*

Training R²: 0.93
Test R²: 0.90
The high R² on both sets indicates that the model generalizes well, minimizing overfitting.

*Output Visualization:*

Figure 1: Scatter plot of actual vs. predicted prices shows a tight clustering around the diagonal, indicating strong predictive accuracy.
Figure 2: A Bar Plot showing summary of test set performance metrics, including RMSE and MAE.

# Q7: What are your assumptions and limitations? What robustness checks did you perform or would you perform?
*Assumptions*
Linear Separability in Transformed Space: Although neural networks can capture non-linear patterns, the dataset must contain enough information for meaningful separation of price categories.
Feature Importance: All selected variables are assumed to have predictive relevance. Features like GarageCars and GrLivArea were identified based on prior knowledge and exploratory data analysis.
No Multicollinearity Among Predictors: Preprocessing removed or combined redundant variables to prevent multicollinearity from impacting model performance.
Data Consistency: The relationships between predictors and the target variable are assumed to be consistent over time and across samples.
*Limitations*
Model Interpretability: Neural networks are inherently less interpretable compared to simpler models (e.g., linear regression). Understanding feature contributions requires additional tools like SHAP or LIME.
Computational Cost: avNNET's lightweight architecture is suitable for small datasets but may struggle with scalability for larger datasets.
Imputation Bias: Imputation methods for missing values (e.g., median for LotFrontage) could introduce bias if missingness is not random.
Limited Generalizability: Results are specific to the Ames, Iowa dataset and may not generalize to other housing markets.
*Robustness Checks*
Cross-Validation: Performed 5-fold cross-validation to evaluate the model's performance across different subsets of the data.
Grid Search for Hyperparameters: Conducted a thorough grid search to optimize parameters like learning rate, hidden neurons, and activation function.
Data Augmentation: Tested the model on bootstrapped samples to ensure performance consistency.
Alternative Architectures: Compared avNNET results with RWeka's neural network implementation to validate findings.
Sensitivity Analysis: Assessed model stability by removing less important variables and retraining.


# Load Libraries

```{r}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE) #Suppress warnings when libraries are loaded.

```

```{r}

library(DDoutlier)
library(ggplot2)
library(gridExtra)
library(tidyverse)
library(RWeka)
library(caret)
library(rminer)
library(tictoc)
library(doParallel)
library(mlbench)

```

# Read in CSV Functions

```{r}

house_prices <- read.csv("C:\\Users\\varun\\Box Sync\\Business Analytics Degree\\Semesters\\Fall Semester 2024\\MKTG 6620\\HW_1\\house-prices-advanced-regression-techniques\\train.csv")

```

# Cleaning the Dataset

This dataset has a few issues which must be addressed through cleaning. Outliers must also be addressed. Afterwards, the dataset can be utilized for modelling.

## Address Missing Values

```{r}

# Define function to check %age of NAs in each column

find_na <- function(x) sum(is.na(x))

# Apply function to each column with map()
missing_values <- map(.x = house_prices, .f = find_na) %>% 
  unlist() %>% 
  data.frame()

missing_values %>% 
  filter(. != 0) %>% 
  mutate(pctg = ./nrow(house_prices)) %>% 
  arrange(desc(pctg)) %>% 
  round(4)

```

### Address Columns greater than 90% of NA's

#### Check Levels for Variables

The below code checks the levels for the variables that have more than 90% of their values missing.

```{r}

unique(house_prices$PoolQC)
unique(house_prices$MiscFeature)
unique(house_prices$Alley)

```

Per the Kaggle Dictionary, Na for `PoolQC` means that there is no pool. Thus, these NA values can be replaced with "No Pool"

Additionally, per the Kaggle Dictionary, NA for `MiscFeature` means that there is no Miscallenous Feature in the house. These NA values can also be replaced with "No Misc Feature". 

Finally NA for `Alley` means that there is no Alley Access for that house. Hence NA for this column can also be replaced with "No Alley". 

#### Recode PoolQC

```{r}

# Recoding cols

# Recode NA in FireplaceQu to "No Fireplace"
house_prices$PoolQC[is.na(house_prices$PoolQC)] <- "No Pool"

# Check that recoding the NA values worked
unique(house_prices$PoolQC)


```

#### Recode MiscFeature

```{r}

# Recode NA in MiscFeature to "No Misc Feature"
house_prices$MiscFeature[is.na(house_prices$MiscFeature)] <- "No Misc Feature"

# Check that recoding the NA values worked
unique(house_prices$PoolQC)

```

#### Recode Alley

```{r}

# Recode NA in MiscFeature to "No Misc Feature"
house_prices$Alley[is.na(house_prices$Alley)] <- "No Alley"

# Check that recoding the NA values worked
unique(house_prices$Alley)

```


## Address Remaining Missing Values 

```{r}

# use above function again to check %age of remaining columns' NA's (<90%)
# recode cols again/impute numeric cols with median and categorical cols with mode

missing_values <- map(.x = house_prices, .f = find_na) %>%  
  unlist() %>% 
  data.frame()

missing_values %>% 
  filter(. != 0) %>% 
  mutate(pctg = ./nrow(house_prices)) %>% 
  arrange(desc(pctg))  

```


#### Analyze Fence Quality

```{r}

unique(house_prices$Fence)

```

The "Fence Quality" is a categorical variable that contains 4 categories: 
* MnPrv
* GdWo
* GdPrv
* MnWw.

Per the Kaggle Competition Website, this column describes the quality of the fence for a property. The Kaggle Website also states that "NA" represents "No Fence".

Thus, NA can be recoded as "No Fence"

```{r}

# Recode NA in Fence to "No Fence"
house_prices$Fence[is.na(house_prices$Fence)] <- "No Fence"

# Check that recoding the NA values worked
unique(house_prices$Fence)

```

Recoding the NA Values was successful. 

#### Analyze Fireplace Qu

`Fireplace Qu` stands for Fireplace Quality and describes the quality of a fireplace. Additionally Kaggle states that NA means that the home does not have a fireplace. 

Thus, any NA values will be recoded to represent "No Fireplace".

```{r}

# Recode NA in FireplaceQu to "No Fireplace"
house_prices$FireplaceQu[is.na(house_prices$FireplaceQu)] <- "No Fireplace"

# Check that recoding the NA values worked
unique(house_prices$FireplaceQu)

```

#### Analyze Lot Frontage

```{r}

# Select only numeric columns
numeric_data <- house_prices %>%
  select(-Id,-OverallQual,-OverallCond,-MSSubClass) %>% 
  select(where(is.numeric))

# Calculate the correlation matrix
cor_matrix <- cor(numeric_data, use = "complete.obs")

# Extract correlations for LotFrontage
lotfrontage_correlations <- cor_matrix["LotFrontage", ]

# Display the correlations
lotfrontage_correlations

```

Lot Frontage does not seem to be strongly correlated with any of the numeric variables. Thus, using a package like Amelia might not be very appropriate. (Amelia would basically create a model that takes the other columns and uses them to predict potential values for the other columns.) 

It should also be noted that although the correlation with the Target Variable `SalePrice` is quite low at .34, it could be still be useful in combination with other variables. 

However in terms of dealing with the NA's, imputation with the median should be fairly reliable. The column is only missing 17.74% of its values. This means that utilizing the median should be reliable since most of the column's values (82.26%) are present. 

```{r}

# Compute the median
median_lot_frontage <- median(house_prices$LotFrontage,na.rm = TRUE)   # median= 69

median_lot_frontage

# Impute with the median
house_prices$LotFrontage[is.na(house_prices$LotFrontage)] <- median_lot_frontage

# Check Missing Values
sum(is.na(house_prices$LotFrontage))

```

#### Analyze all the Garage Variables

The Garage Variables all are missing 81 variables which comprise 5.55% of the dataset: 

* `GarageType`
* `GarageYrBlt`
* `GarageFinish`
* `GarageQual`
* `GarageCond`

According to Kaggle, NA for `GarageType`, `GarageFinish`, `GarageQual`, `GarageCond` represents "No Garage". This means that the NA value can be recoded as "No Garage" for the above columns. 

NA for "Garage Built" however will be a little tricky since including a number like 0 or -9999 or 1 to represent no Garage may confuse any ensemble models. This is because the model may believe that 1 or -9999 is a valid year.

Thus, to avoid any potential confusion or creating possible outliers like 9999, we will just leave NA in `GarageYrBlt` alone. (We can't replace it with a string like "No Garage" because the column type will change and it'll no longer be numeric.) 

Instead, a new column will be created that more explicitly states whether a house has a garage or not. Thus, a new column will be created wtih "0" representing no garage and "1" representing a garage. (This info can be gleaned from the other Garage Columns but we believe that having an explicit column for Garage will make it easier to assess Garage's influence on price and it can be used to compare houses that don't have garages vs having a garage for clustering as well. (We may do clustering and then run a model on that cluster, so having this column will be helpful for that as well.)

```{r}
# create a vector to store all the columns
garage_col <- c("GarageType", "GarageFinish", "GarageQual", "GarageCond")

# Replace NA in specified columns with new value
house_prices <- house_prices %>% 
  mutate(across(all_of(garage_col), ~ replace_na(., "No Garage")))

```

The new column `HasGarage` has been created to represent explicitly whether a house has a garage or not.

```{r}

house_prices <- house_prices %>%
  mutate(HasGarage = ifelse(is.na(GarageYrBlt), "No", "Yes")  # Create a binary indicator for garage presence
  )

```


#### Analyze all the Basement Variables

According to Kaggle, NA for `BsmtExposure`, `BsmtFinType2`, `BsmtQual`, `BsmtCond` and `BsmtFinType1` represents "No Basement". This means that the NA value can be recoded as "No Basement" for the above columns.

They are also all missing roughly the same amount of values in all the columns.

```{r}

# create a vector to store all the columns
basement_col <- c("BsmtExposure", "BsmtFinType2", "BsmtQual", "BsmtCond","BsmtFinType1")

# Replace NA in specified columns with new value
house_prices <- house_prices %>% 
  mutate(across(all_of(basement_col), ~ replace_na(., "No Basement")))

```

#### Analyze all the MasVnr Variables

Kaggle states that if there is no Masonry Veneer for the house, then the value is "None" for the `MasVnrType` variable. This indicates that the variable is genuinely missing values. (i.e any NA values are geniunely missing values and cannot be imputed with "None")

Additionally, if a house does not have a Masonry Veneer, then the `MasVnrArea` is zero. The below code chunk shows this:

```{r}

house_prices %>% 
  select(MasVnrType,MasVnrArea) %>% # Select the following two columns
  filter(MasVnrType == "None") %>% # Only select rows where client marked "None".
  head()

```
The code shows that everytime `MasVnrType` is None, the `MasVnrArea` variable is Zero. Despite this, there are only 8 values missing in both columns which is less than 1% of the data at 0.54% This is a very small amount which means that imputing with median/mode is appropriate.


```{r}

# Define function to find the mode of a character vector
find_mode <- function(x){
  table(x) %>% 
    which.max() %>% 
    names()
}

# Compute with the median
MasVnrType_mode <- find_mode(house_prices$MasVnrType)

MasVnrType_mode

# Impute with the median
house_prices$MasVnrType[is.na(house_prices$MasVnrType)] <- MasVnrType_mode  # mode = "None"

# Check Missing Values
sum(is.na(house_prices$MasVnrType))

```

```{r}

# Compute with the median
median_MasVnrArea <- median(house_prices$MasVnrArea,na.rm = TRUE)  #median = 0
median_MasVnrArea

# Impute with the median
house_prices$MasVnrArea[is.na(house_prices$MasVnrArea)] <- median_MasVnrArea

# Check Missing Values
sum(is.na(house_prices$MasVnrArea))

```

MasVnrType has been successfully imputed with the mode while MasVnrArea has also been successfully imputed with the median.

#### Analyze Electrical Variable

The Electrical Variable is only missing one observation. This missing value can be replaced with the mode since it's a categorical variable.

```{r}
# Get the mode for electrical
electrical_mode <- find_mode(house_prices$Electrical)

# Impute with the mode
house_prices$Electrical[is.na(house_prices$Electrical)] <- electrical_mode

# Check missing values
sum(is.na(house_prices$Electrical))

```

The imputation was successful, there are no missing values, except for `GarageYrBlt` This however was done on purpose, please see "Analyze all the Garage Variables" for more information.

```{r}

map(.x = house_prices, .f = find_na) %>% 
  unlist() %>% 
  data.frame()

```

## Drop ID Column

```{r}

# Drop Id column not needed
house_prices <- house_prices %>% 
  select(-Id)

```

The Id Column is dropped since it is just used to identify each individual house. It does not say anything else or reveal any interesting characteristics about the house.

## Factor Conversion

```{r}

# Factoring all categorical and numeric variables

# Mutate any character variables into factors
house_prices <- house_prices %>% mutate_if(is.character, as.factor)

# Numeric Variables which should be factors stored as a vector
cat_values <- c("OverallQual","OverallCond","MSSubClass", "MoSold")

# Utilize mutate to transform all the columns in the cat_values vector to factors.
house_prices <- house_prices %>% 
  mutate(across(all_of(cat_values), as.factor))

# Display the summary of train_clean
summary(house_prices)


```

This code will all the character variables along with some numeric variables to factors.

* `OverallQual` is being converted because the numeric ratings have a categorical values with them. i.e 10 is "very excellent" while 9 is "excellent". Thus to capture these categorical values, `OverallQual` is being converted to a factor. (This info is from the data dictionary on Kaggle.)
  
  * Overall Quality rates the overall material and finish of the house.

* `OverallCond` also reflect this with the numeric rating containing categorical information. For instance, a 9 is "excellent" while an 8 is "very good". (This info is from the data dictionary on Kaggle.)

  * Overall Cond rates the overall condition of the house.

* `MSSubclass` also contains categorical information which has been represented as a numeric variable. Thus, to capture this information, this variable will also be converted to a factor. For instance, "20" represents a "1-STORY 1946 & NEWER ALL STYLES"

  * Identifies the type of dwelling involved in a sale.
  
* `MoSold` just represents the month that a house was sold. This should be a categorical variable, it makes no sense to calculate the "median" month, etc. There additionally is no meaningful numeric diference between "January" and "February" for instance. These are just names of the months which this variable represnts. It has just been created as a numeric column.

## Outliers Analysis

To remove the outliers, the LOF (Local Outlier Factor) algorithm will be utilized. This approach utilizes LOOP (Local Outlier Probability) where it utilizes the LOF Scores to calculate the probability of an observation being an outlier.

Afterwards any observations that have a greater than 75% probability of being an outlier will be removed from the dataset.

```{r}

numeric_columns <- sapply(house_prices, is.numeric) 

house_clean_numeric <- house_prices[, numeric_columns] # subset train_clean using only numeric cols  (33 cols)

house_clean_numeric <- house_clean_numeric %>% select(-GarageYrBlt)
# Can't have missing values

lof<- sapply(5:10, function(k) LOF(house_clean_numeric,k))

```

This above code calculates 6 LOF Values for each row in the dataset. For instance, for row 1, the chunk will calculate the LOF for 5 neighbors, 6 neighbors, etc. up to 10 neighbors.

```{r}

#calcuate median lof of different neighborhood sizes.

lof<-apply(lof,1,median)

```

The above code then calculates the median LOF score for each row from the 6 LOF scores previously generated for each row.

```{r}

loop<- sapply(5:10, function(k) LOOP(house_clean_numeric,k))
loop<-apply(loop,1,median)

# Remove observation with LOOP more than .75
# hpc = house_prices_clean
hpc <- house_prices[loop <= 0.75, ]

```

A potential issue however that was present in the last approach with utilizing the LOF Algorithm was that the threshold was a little arbitrarily decided. A potential work around to this issue is to utilize LOOP. LOOP utilizes LOF Scores to calculate the probability of an observation being an outlier. Afterwards, any LOOP values that are greater than .75 would be filtered out from the dataset.

This method is more robust then the LOF Approach #1 where a threshold was somewhat arbitrarily selected based on some guidelines. Additionally, this method calculates the median LOF Score across different number of neighbors which gives a more robust view of an observation's Local Reachabilty Density and LOF score. (This is because we are taking the median of several LOF Scores as opposed to just one LOF score like in the previous method.)

There is the issue of selecting which probabilty threshold should be selected for determining if something is an outlier. However, the median LOF Scores which are used to calculate the LOOP should provide a better view than just using LOF Scores from a fixed neighbor number. (i.e calculating the LOF Scores from just 6 neighbors for every observation.)

```{r}

row_reduction_02 <- ((1460 - nrow(hpc))/1460) * 100

round(row_reduction_02,2)


nrow(hpc)
ncol(hpc)

```

The dataset now has 1,439 rows and 81 columns. This is a 1.44% reduction which means that 98.56% of the data has been conserved.

This approach is more robust approach because it considers the impact of multiple neighbors (5 nearest neighbors to 10 nearest neighbors) by calculating the median. Calculating the median of the different LOF scores helps to show a more consistent picture of an observation's density point. 

Furthermore, when combined with LOOP, this method can avoid arbitrary thresholds as previously explained. Deciding on a probability however is still a little arbitrary, but a bit more easier than deciding a threshold score which is dependent on the LOF Score which depends on a an observation's local reachable density point.

# Neural Networks

```{r}

# Designate a shortened name MLP for the MultilayerPercentron ANN method in RWeka
MLP <- make_Weka_classifier("weka/classifiers/functions/MultilayerPerceptron")

```

## Partition Dataset

```{r}

# Partititon Dataset via 80/20 split
set.seed(123)
row_indices <- createDataPartition(hpc$SalePrice,p=0.8,list=FALSE)

# Assign rows to train and test set respectively 
train_set <- hpc[row_indices,] # 80% of data should be in train
test_set <- hpc[-row_indices,] # 20 % of data should be in test

nrow(train_set) # Display Number of rows
nrow(test_set) # Display Number of rows

total_rows <- nrow(train_set) + nrow(test_set)
total_rows
```

The partition is successful, the number of rows from the train set (1153) and the test set(286) is equal to 1439 rows.

# Set up Cores 

```{r}

num_cores <- detectCores()
num_cores <- num_cores - 1 


```



# First Model

```{r}

tic()
doParallel::registerDoParallel(cores = num_cores) 
ANN_00 <- MLP(SalePrice ~ . , data = train_set)
toc()

#saveRDS(ANN_00, file = "ANN_00.rds")


```
## Generate Predictions and Metrics

```{r}
# Generate Evaluation Metrics
metrics_list <- c("MAE","RMSE","MAPE","RMSPE","RAE","RRSE","R2")

# Generate Predictions
#tr = train
ANN_00_train <- predict(ANN_00,newdata = train_set)
# te = test
ANN_00_test <- predict(ANN_00, test_set)

# Rounding Function for the metrics 
round_metrics <- function(x) {
  return(lapply(x, function(y) {
    if (is.numeric(y)) {
      return(round(y, 2))
    } else {
      return(y)
    }
  }))
}


# In the Weka model, there are default paramters which are list down below
#l <- 0.3 - learn rate
#m <- 0.2 - momentum
#n <-500 - number of epochs/passes
#h <- 'a' - #The hidden nodes to be created on each layer:
  # an integer, or the letters 'a' = (# of attribs + # of classes) / 2, 
  #'i' = # of attribs, 'o' = # of classes, 't' = (# of attribs + # of classes)
  # default of H is 'a'.

```

### Train Metrics

```{r}
ANN_00_tr_met <- mmetric(train_set$SalePrice,ANN_00_train,metric = metrics_list)
round_metrics(ANN_00_tr_met)

```

The model seems to be doing good on the train set with an $R^2$ of 93%. This means that the model is able to explain 93% of the variation within the data.

The RMSE also seems fairly decent at 23190.96, dollars. This means that the model's predictions are of on average by 23190.96 dollars.

### Test Metrics

```{r}

ANN_00_te_met <- mmetric(test_set$SalePrice,ANN_00_test,metric = metrics_list)
round_metrics(ANN_00_te_met)

```

It appears that the model is over-fitting when it's cross-validated with the test set. For instance, $R^2$ sees a modest decrease of 10% from the train to the test set. (train_set $R^2$: 93%, test_set$R^2$: 83%)

The RMSE however sees a large increase of 8150.59 dollars from the train to the test sets. (train_set RMSE: 23190.96, test_set RMSE: 31341.55)

This means that the model is picking up on noise in the data and is very complicated. A potential reason could be that the model has 171 hidden nodes on 1 layer.

# Second Model

```{r}

# These are the default values for MLP, except for h
l <- 0.3
m <- 0.2
n <-500
h <- '86' # Split the number of hidden nodes by half (171/2)
# 86 means 1 layer with 86 nodes

tic()
doParallel::registerDoParallel(cores = num_cores) 
ANN_02 <- MLP(SalePrice ~ . , data = train_set,control= Weka_control(L = l, M = m, N = n, H = h))
toc()

saveRDS(ANN_02, file = "ANN_02.rds")

summary(ANN_02)


```

Since the first model slightly overfit, we'll attempt to simplify this model by reducing the number of nodes on the hidden layer. We can do this by cutting the number of nodes in half. If one looks at the first, model there were 171 nodes, so we'll cut this in half and train the model. 

## Generate Predictions and Train Metrics

```{r}

ANN_02_train <- predict(ANN_02,newdata = train_set)
ANN_02_test <- predict(ANN_02,newdata= test_set)

ANN_02_tr_met <- mmetric(train_set$SalePrice,ANN_02_train,metric = metrics_list)
round_metrics(ANN_02_tr_met)


```

## Test Metrics

```{r}

ANN_02_te_met <- mmetric(test_set$SalePrice,ANN_02_test,metric = metrics_list)
round_metrics(ANN_02_te_met)

```

Reducing the nodes by half on the hidden layer has caused the model overfit even more severely.

This model's RMSE is: 
* train_set RMSE: 42904.12
* test_set RMSE: 55100.95
* difference: 12196.83
 
The RMSE values are much higher than the first model's RMSE Values which are listed down below for reference.

First model's RMSE: 
* train_set RMSE: 23190.96
* test_set RMSE: 31341.55
* difference: 8150.59

Additionally the $R^2$ on the first model was: 
* train $R^2$: 93%
* test $R^2$: 83%
* difference: 10%

$R^2$ on second model for train and test:
* train $R^2$: 93%
* test $R^2$: 78%
* difference: 15% 

These metrics indicate that the number of nodes should not be reduced so drastically since it causes a large decline in model performance. The model actually overfits more.

The third model will thus seek to reduce the number of nodes but not by much this time around.

# Third Model

```{r}

# These are the default values for MLP, except for h
l <- 0.3
m <- 0.2
n <-500
h <- '128' # Split the number of hidden nodes by half

tic()
ANN_03 <- MLP(SalePrice ~ . , data = train_set,control= Weka_control(L = l, M = m, N = n, H = h))
toc()

saveRDS(ANN_03, file = "ANN_03.rds")

```

# Generate Predictions and Train Metrics

```{r}

ANN_03_train <- predict(ANN_03,newdata = train_set)
ANN_03_test <- predict(ANN_03,newdata= test_set)

ANN_03_tr_met <- mmetric(train_set$SalePrice,ANN_03_train,metric = metrics_list)
round_metrics(ANN_03_tr_met)

```

# Test Metrics

```{r}

ANN_03_te_met <- mmetric(test_set$SalePrice,ANN_03_test,metric = metrics_list)
round_metrics(ANN_03_te_met)

```

This model more severely overfits when compared to the previous two models.  

This Models Train and Test RMSE (M3), $R^2$: 

* Train RMSE: 14859.19
* Test RMSE: 41730.83
* Difference between RMSE: 26871.64
* Train $R^2$: 97%
* Train $R^2$: 78%
* Difference between $R^2$: 19% 

Previous Model's Train and Test RMSE (M2), $R^2$:
* Train RMSE: 78891.49 dollars 
* Test RMSE: 72526.44
* Difference between RMSE: 26871.64
* Train $R^2$: 93%
* Test $R^2$: 78%
* Difference between $R^2$: 15% 

First Model's Train and Test RMSE (M1), $R^2$:
* Train RMSE: 23190.96
* Test RMSE: 31341.55
* Difference between RMSE: 8150.59
* Train $R^2$: 93%
* Test $R^2$: 83%
* Difference between $R^2$: 10% 

The Third Model's RMSE is much lower in terms of absolute values however the difference between the train and test sets for RMSE is the same as the RMSE difference between the train and test sets for the second model.

Additionally, it has the largest decrease in $R^2$ of 19% when compared to the other models. (M1 has a decrease of 10% while M2 has a decrease of 15%.)

# Fourth Model

It seems that reducing the nodes does not do much to help with overfitting. In fact, in some cases it actually increases the severity of the overfitting like with the 3rd model. This can be seen through the drastic decrease of $R^2$. Additionally, with the second model (massive node reduction), the RMSE increase drastically when compared to the other 2 models. (This is because less nodes means that model is unable to pick up more complex patterns which results in the model's predictions being further off.)

It should however be noted that the first model was only over fitting slightly with a decrease of 10% from train to test $R^2$. The RMSE for the first model for the test sets was also lower in absolute terms when compared to the other 2 models. Thus, another parameter should be experimented with instead like regularization. 

RWeka does not directly have a regulariztion parameter but this can be emulated through the learn rate and momentum. Keeping a slower learn rate will make the model learn more 

```{r}

# These are the default values for MLP, except for h and l
l <- 0.1
m <- 0.5
n <-500
h <- '176' #

tic()
doParallel::registerDoParallel(cores = num_cores) 
ANN_04 <- MLP(SalePrice ~ . , data = train_set,control= Weka_control(L = l, M = m, N = n, H = h))
toc()

```

## Generate Predictions and Train Metrics

```{r}

ANN_04_train_pred <- predict(ANN_04, newdata = train_set)
ANN_04_test_pred <- predict(ANN_04, newdata = test_set)

ANN_04_tr_met <- mmetric(train_set$SalePrice,ANN_04_train_pred,metric = metrics_list)
round_metrics(ANN_04_tr_met)


```

## Test Metrics

```{r}

ANN_04_te_met <- mmetric(test_set$SalePrice,ANN_04_test_pred,metric = metrics_list)

round_metrics(ANN_04_te_met)

```


# Grid Search



## AVNET

We'll try AVNET with the decay parameters to see if the model performance can be improved. 

```{r}

train_set <- train_set[, !colnames(train_set) %in% c("GarageYrBlt")]
test_set <- test_set[, !colnames(test_set) %in% c("GarageYrBlt")]

train_predictor <- train_set[,-79] # SalePrice is col 80 in data-set
train_outcome <- train_set [,79] 

test_predictor <- test_set[,-79]
test_outcome <- test_set[,79]


ctrl <- trainControl(method = "repeatedcv",   # cross-validation
                     number = 5,               # 5 folds
                     repeats = 5,              # 5 repetitions
                     savePredictions = "all"   # Save all predictions to help with diagnostics
)

avnnetGrid <- expand.grid(decay = c(.001,.005),
size = c(1:6),
bag = FALSE
)

tic()
doParallel::registerDoParallel(cores = num_cores) 
avnnet_fit <- train(train_predictor, train_outcome,
method = "avNNet",
tuneGrid = avnnetGrid,
trControl = ctrl,
#preProc = c("center", "scale"),
trace = FALSE,
metric= "RMSE",
allowParallel=TRUE,
maxit = 100)
toc()


```

```{r}

plot(avnnet_fit)

predicted_outcome <- predict(avnnet_fit, newdata = test_predictor)
# Evaluate performance
metrics <- postResample(pred = predicted_outcome, obs = test_outcome)

# Print the metrics
print(metrics)

```

## Grid Search

It's quite apparent based of the previous neural networks that there are several different hyper-parameters from the Learn and Momentum Rates to the number of hidden nodes. It would take an incredibly long time to experiment with all of these hyper-paramters to find a model that is generalizable. Some of the hyper-paramters cause the model to severely underfit while other hyper-paramters cause the model to overfit. 

It's also very computationally expensive to experiment and select the model with the best hyper-parameters. Thus a more efficient method is required to find a model that generalizes well to new data and does not overfit or underfit.

This method is Grid Search which we can use to find a model that generalizes well to new data. Grid-Search will search through various hyperparameter combinations, till it finds the best model based on a specified metric.

```{r}


train_set <- train_set[, !colnames(train_set) %in% c("GarageYrBlt")]
test_set <- test_set[, !colnames(test_set) %in% c("GarageYrBlt")]

train_predictor <- train_set[,-79] # SalePrice is col 80 in data-set
train_outcome <- train_set [,79] 

test_predictor <- test_set[,-79]
test_outcome <- test_set[,79]
# Define parameter grid
learning_rates <- c(0.1, 0.01, 0.001)
momentums <- c(0.2, 0.5, 0.9)
hidden_layers <- c("5", "10,5", "20","32,64,83")  # Different layer configurations
epochs <- c(100, 200)

# Placeholder for results
results <- data.frame(LearningRate = numeric(),
                      Momentum = numeric(),
                      HiddenLayers = character(),
                      Epochs = numeric(),
                      RMSE = numeric(),
                      stringsAsFactors = TRUE)

# Grid search
tic()
set.seed(123)  # Use this seed for consistency in future work
for (lr in learning_rates) {
  for (mom in momentums) {
    for (hl in hidden_layers) {
      for (epoch in epochs) {
        
        # Train model with current parameters
        model <- MLP(train_outcome ~ ., 
                     data = train_predictor, 
                     control = Weka_control(L = lr, M = mom, H = hl, N = epoch))
        
        # Evaluate performance (e.g., RMSE)
        predictions <- predict(model, test_set)
        rmse <- sqrt(mean((predictions - test_outcome)^2))
        
        # Store results
        results <- rbind(results, data.frame(
          LearningRate = lr,
          Momentum = mom,
          HiddenLayers = hl,
          Epochs = epoch,
          RMSE = rmse
        ))
      }
    }
  }
}

# Find best parameters
best_model <- results[which.min(results$RMSE), ]
print(best_model)
toc()

# My old HP G10 Revolve Notebook Grid Search Train Time
# Approximately 38 minutes, so maybe 37.87 min

# My Amma's Dell Laptop Grid Search Train Time
# Approximately 15 minutes, so maybe 14.75 min

# My new HP Victus Grid Search Train Time:
# Recorded right now, so I can say the exact time: 11.7425 minutes which rounds up to 12 minutes. Slightly faster than my mom's laptop.

# It looks like R is utilizing the Nvidia RTX GeForce 3050 chip when training the model, but not to its full extent. Even then though, it's still much faster than my old laptop and slightly faster than my mom's laptop.

# To get the true running time, we would have to run this grid search on all 3 machines multiple times.

```

```{r}

# Extract best parameters
best_lr <- best_model$LearningRate
best_mom <- best_model$Momentum
best_hl <- best_model$HiddenLayers
best_epochs <- best_model$Epochs

# Train the model with the best parameters
final_model <- MLP(train_outcome ~ ., 
                   data = train_set, 
                   control = Weka_control(L = best_lr, M = best_mom, H = best_hl, N = best_epochs))

final_predictions <- predict(final_model, test_set)

# Calculate RMSE
rmse <- sqrt(mean((final_predictions - test_outcome)^2))

# Calculate R-squared
sse <- sum((final_predictions - test_outcome)^2)  # Sum of Squared Errors
sst <- sum((test_outcome - mean(test_outcome))^2) # Total Sum of Squares
r_squared <- 1 - (sse / sst)

# Calculate MAE
mae <- mean(abs(final_predictions - test_outcome))

# Print metrics
cat("RMSE:", rmse, "\n")
cat("R-squared:", r_squared, "\n")
cat("MAE:", mae, "\n")

```

## Train Set

```{r}

train_predictions <- predict(final_model, train_set)

# Calculate RMSE
train_rmse <- sqrt(mean((train_predictions - train_outcome)^2))

# Calculate R-squared
train_sse <- sum((train_predictions - train_outcome)^2)  # Sum of Squared Errors
train_sst <- sum((train_outcome - mean(train_outcome))^2) # Total Sum of Squares
train_r_squared <- 1 - (train_sse / train_sst)

# Calculate MAE
train_mae <- mean(abs(train_predictions - train_outcome))

# Print metrics
cat("Training Set Metrics:\n")
cat("RMSE:", train_rmse, "\n")
cat("R-squared:", train_r_squared, "\n")
cat("MAE:", train_mae, "\n")

```

```{r}

comparison_metrics <- data.frame(
  Dataset = c("Train", "Test"),
  RMSE = c(train_rmse, rmse),
  Rsquared = c(train_r_squared, r_squared),
  MAE = c(train_mae, mae)
)

print(comparison_metrics)

```

# Visualizations

## Scatter Plot: Actual vs. Predicted Prices (Test Set)

```{r}

# Scatter Plot: Actual vs. Predicted Prices (Test Set)
scatter_plot <- ggplot(data = data.frame(Actual = test_outcome, Predicted = final_predictions), aes(x = Actual, y = Predicted)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Scatter Plot: Actual vs. Predicted Prices (Test Set)",
       x = "Actual Sale Prices",
       y = "Predicted Sale Prices") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

plot(scatter_plot)

```

Scatter Plot:

It Plots test_outcome (actual values) vs. final_predictions (predicted values). Visual feedback on how well the model's predictions align with actual values. A tight clustering around the diagonal line indicates a good fit.
The diagonal red dashed line indicates perfect predictions. Points close to this line signify accurate predictions.

## Comparison Table

```{r}

# Comparison Table Visualization

comparison_metrics$Dataset <- factor(comparison_metrics$Dataset, levels = c("Train", "Test"))
comparison_plot <- ggplot(comparison_metrics, aes(x = Dataset)) +
  geom_col(aes(y = RMSE, fill = Dataset), position = position_dodge(width = 0.8), alpha = 0.8) +
  geom_text(aes(y = RMSE, label = round(RMSE, 2)), vjust = -0.5, color = "black", size = 4) +
  labs(title = "Comparison of RMSE by Dataset",
       x = "Dataset",
       y = "RMSE") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  scale_fill_manual(values = c("Train" = "skyblue", "Test" = "orange"))

plot(comparison_plot)

```

Comparison Table Visualization:

Displays the RMSE values for the train and test datasets as bar heights. It is an intuitive comparison of RMSE for training and testing datasets to check for overfitting or underfitting.
It also annotates the bar heights with their corresponding RMSE values for clarity.


