---
title: "Group Project"
author: "Varun Selvam, Ahsan Ahmad, Richard Lim"
date: "2023-11-26"
output: 
  html_document:
    theme: flatly
    df_print: paged
    toc: true
    toc_float:
      smooth_scroll: true
---
```{r setup, include=FALSE}
# Suppress warnings libraries
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE) 
options(scipen = 2)
```

## Introduction & Project goal

### Introduction

In this group project, we are expecting to build the maximum predictive accuracy by using all available predictors & interactions.there is no restriction on variables to utilize, or on combinations of variables—interactions and polynomial terms are encouraged, as appropriate. 

### Project goal

This group project is aimed to develop a linear model predicting housing prices in Ames, IA using dataset provided in **Kaggle.com**.  Model performance benchmark for minimum estimated out-of-sample $R^2$ is 0.85.

## Description of the data

### Description
There are two datasets for the group project. One is `train.csv` and the other is `test.csv` which are with 1460 and 1459 observations, respectively. Data includes the categorical & numeric features of properties (Predictors) and their corresponding `SaePrice`. (Target)


### Data Profile
* SalePrice - the property's sale price in dollars. This is the target variable that you're trying to predict.
* MSSubClass: The building class
* MSZoning: The general zoning classification
* LotFrontage: Linear feet of street connected to property
* LotArea: Lot size in square feet
* Street: Type of road access
* Alley: Type of alley access
* LotShape: General shape of property
* LandContour: Flatness of the property
* Utilities: Type of utilities available
* LotConfig: Lot configuration
* LandSlope: Slope of property
* Neighborhood: Physical locations within Ames city limits
* Condition1: Proximity to main road or railroad
* Condition2: Proximity to main road or railroad (if a second is present)
* BldgType: Type of dwelling
* HouseStyle: Style of dwelling
* OverallQual: Overall material and finish quality
* OverallCond: Overall condition rating
* YearBuilt: Original construction date
* YearRemodAdd: Remodel date
* RoofStyle: Type of roof
* RoofMatl: Roof material
* Exterior1st: Exterior covering on house
* Exterior2nd: Exterior covering on house (if more than one material)
* MasVnrType: Masonry veneer type
* MasVnrArea: Masonry veneer area in square feet
* ExterQual: Exterior material quality
* ExterCond: Present condition of the material on the exterior
* Foundation: Type of foundation
* BsmtQual: Height of the basement
* BsmtCond: General condition of the basement
* BsmtExposure: Walkout or garden level basement walls
* BsmtFinType1: Quality of basement finished area
* BsmtFinSF1: Type 1 finished square feet
* BsmtFinType2: Quality of second finished area (if present)
* BsmtFinSF2: Type 2 finished square feet
* BsmtUnfSF: Unfinished square feet of basement area
* TotalBsmtSF: Total square feet of basement area
* Heating: Type of heating
* HeatingQC: Heating quality and condition
* CentralAir: Central air conditioning
* Electrical: Electrical system
* 1stFlrSF: First Floor square feet
* 2ndFlrSF: Second floor square feet
* LowQualFinSF: Low quality finished square feet (all floors)
* GrLivArea: Above grade (ground) living area square feet
* BsmtFullBath: Basement full bathrooms
* BsmtHalfBath: Basement half bathrooms
* FullBath: Full bathrooms above grade
* HalfBath: Half baths above grade
* Bedroom: Number of bedrooms above basement level
* Kitchen: Number of kitchens
* KitchenQual: Kitchen quality
* TotRmsAbvGrd: Total rooms above grade (does not include bathrooms)
* Functional: Home functionality rating
* Fireplaces: Number of fireplaces
* FireplaceQu: Fireplace quality
* GarageType: Garage location
* GarageYrBlt: Year garage was built
* GarageFinish: Interior finish of the garage
* GarageCars: Size of garage in car capacity
* GarageArea: Size of garage in square feet
* GarageQual: Garage quality
* GarageCond: Garage condition
* PavedDrive: Paved driveway
* WoodDeckSF: Wood deck area in square feet
* OpenPorchSF: Open porch area in square feet
* EnclosedPorch: Enclosed porch area in square feet
* 3SsnPorch: Three season porch area in square feet
* ScreenPorch: Screen porch area in square feet
* PoolArea: Pool area in square feet
* PoolQC: Pool quality
* Fence: Fence quality
* MiscFeature: Miscellaneous feature not covered in other categories
* MiscVal: $Value of miscellaneous feature
* MoSold: Month Sold
* YrSold: Year Sold
* SaleType: Type of sale
* SaleCondition: Condition of sale

## Load Libraries

```{r}
# load libraries for project
library(dplyr) 
library(ggplot2)
library(tidyr)
library(rpart)
library(rpart.plot)
```

## Load dataset

```{r}
# load Data sets
train <- read.csv("train.csv")
test <- read.csv("test.csv")
```

```{r}
# 5 sameples of train.csv
head(train)
```

Display the first 6 values of the Train Set, to gain a perspective of the data and its variables.

```{r}
# 5 sameples of test.csv
head(test)
```

Display the first 6 values of the Train Set, to gain a perspective of the data and its variables.

## Selecting Predictors

##Selecting Predictors (Part 1)
```{r}
#Part 1: Classification Tree: 

train_tree <- train %>% #Assign changes to train data set to a new variable
  select(-Id) #Remove ID as a predictor, it's not needed.
  
house_prices_tree <- rpart(formula = SalePrice ~ ., data = train_tree) 
# Assign tree to a variable titled "house_prices_tree".
house_prices_tree #Display Tree

```

Based on this classification tree, the top predictors are: 

- Overall Quality
- Neighborhood
- First Floor Sqft 
- Ground Living Area
- Basement Finished Sqft 1 

##Selecting Predictors (Part 1)
```{r}
# Removing all predictors with even one NA value and saving it to a dataset

without_na_train <-
  train %>% 
  select(names(which(colSums(is.na(.)) == 0)))

# Creating a Multiple Linear Regression Model with all the remaining predictors
model_a <- lm(SalePrice ~ ., data = without_na_train)
model_a_summary <- summary(model_a)

# Creating a Dataframe with Predictor/Variable name and it's corresponding P-Value and arranging it in ascending order to get the top 5 predictors
p_values <- model_a_summary$coefficients[, "Pr(>|t|)"]
coefficients_p_value <- data.frame(
  Variable_name = rownames(model_a_summary$coefficients),
  P_Value = p_values
)
coefficients_p_value %>% 
  arrange(P_Value) %>% 
  head(n = 10)

view(p_values)

without_na_train %>% 
  select(RoofMatl, `1stFlrSF`, `2ndFlrSF`, BsmtFinSF1 , OverallQual, Neighborhood, Utilities, YearBuilt)
```
The above list displays the predictors that the have the lowest p-values. In this context, this means that the values observed for these variables 

The predictors were selected based on the Classification Tree and their p-values. Predictors with the lowest p-values were selected.


## Data Missing Handling

```{r}
count_missings <- function(x) sum(is.na(x)) # Sum provides a total of how many values are missing for each variable. 

# Summarizes all the missing values for each variable.
train %>% # Data set name
  summarize_all(count_missings) %>% #summarize missing values for any of the predictors.
  t()

# Summarize any missing values for the specific predictors that will be utilized in the model. 
train %>% #Dat Set name
  select(OverallQual, Neighborhood, X1stFlrSF, GrLivArea, BsmtFinSF1, RoofMatl, SaleType, Utilities, Condition1, YearBuilt, X2ndFlrSF, SaleType, KitchenQual, LotArea, ExterQual, OverallCond, BsmtFinSF2 ) %>% # Predictors that this model will utilize.
  summarize_all(count_missings) # Summarize all missing values for any of the predictors.

```
Based on the output above, there some variables that have several values missing, such as Alley with 1369 values and LotFrontage at 259 values. 

We have however created a table that shows if there are any missing values for the specific predictors that we have chosen. Based of that table, there are no missing values for those selected predictors.

## Data Cleansing

```{r}
ggplot(data = train, mapping = aes(x = SalePrice)) +
  geom_histogram(fill = 'orange') +
  labs(title = "Histogram of Sale Price") +
  theme_classic()



train %>% 
  ggplot(aes(sqft, price, col = city))+ # Color the lines in the plot by city
  geom_smooth(method = "lm", se = F) + #Set method to LR
  labs(title = "Price ~ sqft * city") # Give title to plot
```
The Histogram for Sale Price is right skewed, however much of the data still appears to near the center.
```{r}
# overwrite original dataset with the new changes
train <- train %>% 
  mutate(OverallQual = factor(OverallQual)) %>% # Change overall quality to a factor.
  mutate(Neighborhood = factor(Neighborhood)) %>% # Change Neighborhood to a factor.
  mutate(RoofMatl = factor(RoofMatl)) %>% # Change Roof Material to a factor.
  mutate(SaleType = factor(SaleType)) %>% # Change Sale Type to a factor.
  mutate(Utilities = factor(Utilities)) %>% # Change Utilities to a factor.
  mutate(Condition1 = factor(Condition1)) %>% # Change Condition1 to a factor.
  mutate(KitchenQual = factor(KitchenQual)) %>% # Change Kitchen Quality to a factor.
  mutate(ExterQual = factor(ExterQual)) %>% # Change External Quality to a factor.
  mutate(OverallCond = factor(OverallCond)) # Change Overall Condition to a factor.
str(train)
```

## Data split for validation

```{r}
#Set seed for reproducibility
set.seed(100) 
index <- sample(x = 1:nrow(train), size = nrow(train)*.7, replace = F) # Randomly sample 70% of the rows

head(index) # Display row numbers.

# Subset train using the index and assign it to train_fold.
train_fold <- train[index, ]

# Subset all remaining rows into the validation fold.
validation_fold <- train[-index, ]

```


## Data Modeling

```{r}
SalePrice_Model <- lm(SalePrice ~ OverallQual + YearBuilt + Neighborhood * X1stFlrSF + GrLivArea + BsmtFinSF1 + RoofMatl + Neighborhood * X2ndFlrSF + SaleType + Utilities + Condition1 + KitchenQual + LotArea + ExterQual + OverallCond + BsmtFinSF2, data = train_fold) #Data must be train-fold because we are training the model, so train_fold data should be used.

summary(SalePrice_Model) #Provide summary of the Sale Price Model.
```

```{r}
rmse <- function(observed, predicted) sqrt(mean((observed - predicted)^2)) #Create RMSE function for test data

R2 <- function(observed, predicted){ #Create R-Squared function for test data
  TSS <- sum((observed - mean(observed))^2)
  RSS <- sum((observed - predicted)^2)
  1- RSS/TSS
}

# Calculate Estimated RMSE
rmse(observed = train_fold$SalePrice, #observed is the SalePrice values that are already in the dataset. 
     predicted = fitted(SalePrice_Model)) #fitted represents the values that the model predicted.
R2(train_fold$SalePrice, fitted(SalePrice_Model))
```

## Data model performance for validation

```{r}
predictions <- predict(SalePrice_Model, newdata = validation_fold)

# Create functions for calculating RMSE and R-squared (necessary for estimating Out of sample performance)

rmse(validation_fold$SalePrice, predictions)
R2(validation_fold$SalePrice, predictions)
```


## Data Engineering for test dataset

```{r}
test <- test %>%  #Override test dataset with changes to make test dataset match the train dataset.
  mutate(OverallQual = factor(OverallQual)) %>% # Change overall quality to a factor.
  mutate(Neighborhood = factor(Neighborhood)) %>% 
  mutate(RoofMatl = factor(RoofMatl)) %>% 
  mutate(SaleType = factor(SaleType)) %>% 
  mutate(Utilities = factor(Utilities)) %>%
  mutate(Condition1 = factor(Condition1)) %>% 
  mutate(KitchenQual = factor(KitchenQual)) %>% 
  mutate(ExterQual = factor(ExterQual)) %>%
  mutate(OverallCond = factor(OverallCond))

# 3. Checking for NA Values for our five selected predictors in the test set.
count_missings <- function(x) sum(is.na(x))

test %>% 
  select(RoofMatl, X1stFlrSF, X2ndFlrSF, BsmtFinSF1 , OverallQual, Utilities, SaleType, GrLivArea, Neighborhood, Condition1, KitchenQual, LotArea, ExterQual, OverallCond, BsmtFinSF2) %>% 
  summarize_all(count_missings) 

```

```{r}
ggplot(data = train, mapping = aes(x = RoofMatl, y = SalePrice)) + 
  geom_boxplot()
```



```{r}
# Replacing the missing value for BsmtFinSF1 with the median

test$BsmtFinSF1 <- as.numeric((test$BsmtFinSF1))
test$BsmtFinSF2 <- as.numeric((test$BsmtFinSF2))

test <- test %>% 
  mutate(BsmtFinSF1 = replace_na(data = BsmtFinSF1, replace = median(BsmtFinSF1, na.rm = TRUE))) %>% 
  mutate(Utilities = replace_na(data = Utilities, replace = "AllPub")) %>% 
  mutate(SaleType = replace_na(data = SaleType, replace = "New")) %>%
  mutate(KitchenQual = replace_na(data = KitchenQual, replace = "Gd")) %>% 
  mutate(BsmtFinSF2 = replace_na(data = BsmtFinSF2, replace = median(BsmtFinSF1, na.rm = TRUE)))
```


## Kaggle Submission

```{r}
# Assign a new linear regression model with the same predictors to a new variable. Use the full data set this time.
kaggle_model <- lm(SalePrice ~ OverallQual + YearBuilt + Neighborhood * X1stFlrSF + GrLivArea + BsmtFinSF1 + RoofMatl + Neighborhood * X2ndFlrSF + SaleType + Utilities + Condition1 + KitchenQual + LotArea + ExterQual + OverallCond + BsmtFinSF2, data = train) 

# Utilize newdata argument.
kaggle_predictions <- predict(kaggle_model, newdata = test) 

#Object for predict is the kaggle_model which is the regression model that is using the full train set.

# see first few predictions
head(kaggle_predictions)

kaggle_submission <- test %>% #Assign changes from test to new variable called kaggle_submission
  select(Id) %>% #Match Kaggle Competition format rules by selecting ID
  mutate(SalePrice = kaggle_predictions) #Create new column in data set that contains the predictions for the Sale Price values.

# Check
head(kaggle_submission) #Check that formatting is correct.

# write to csv
write.csv(kaggle_submission, "kaggle_submission_14.csv", row.names = F)
```

## Kaggle Rank & Score

- Rank: 2767
- Score: 0.14910
